---
title: "A map of the US election results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Upload libraries:
  
```{r libr, message = FALSE}
rm(list = ls(all=T)) #clear workspace
library(dplyr)
library(readr)
library(stringr)
library(tidyr)
library(readxl)
library(classInt)
library(RColorBrewer)
library(maptools) #to read shapefiles
```
2. Download the data files (note they are not ready for use but need some cleaning as there are more areas in the excel files than polygons in the shape file). I copy here the code as I have used it in my script but it's available at <a href="https://rpubs.com/dgrtwo/county-results>RPubs</a> thanks to David Robinson.

```{r download data, message = FALSE}
download.file("http://www2.census.gov/prod2/statcomp/usac/excel/LND01.xls", "LND01.xls")
download.file("http://www2.census.gov/prod2/statcomp/usac/excel/POP01.xls", "POP01.xls")
```
according to metadata, this is Land Area in 2010 and resident population in 2010:
```{r us_county, message = TRUE}
us_county_area <- read_excel("LND01.xls") %>%
  transmute(CountyCode = as.character(as.integer(STCOU)), Area = LND110210D)

us_county_population <- read_excel("POP01.xls") %>%
  transmute(CountyCode = as.character(as.integer(STCOU)),Population = POP010210D)
```
3. Adjust data
```{r adjust data, message = FALSE}
election_url <- "https://raw.githubusercontent.com/Prooffreader/election_2016_data/master/data/presidential_general_election_2016_by_county.csv"
county_data <- read_csv(election_url) %>%
  group_by(CountyCode = as.character(fips)) %>%
  mutate(TotalVotes = sum(votes)) %>%
  ungroup() %>%
  mutate(name = str_replace(name, ".\\. ", "")) %>%
  filter(name %in% c("Trump", "Clinton", "Johnson", "Stein")) %>%
  transmute(County = str_replace(geo_name, " County", ""),
            State = state,
            CountyCode = as.character(fips),
            Candidate = name,
            Percent = vote_pct / 100,
            TotalVotes) %>%
  spread(Candidate, Percent, fill = 0) %>%
  inner_join(us_county_population, by = "CountyCode") %>%
  inner_join(us_county_area, by = "CountyCode")
```
you can save the data into a csv file:
```{r libraries, message = FALSE}
# write_csv(county_data, "county_election_2016.csv")
```
4. Upload data and shape files

```{r shapefileupload, message = FALSE}
setwd("/Users/ac1y15/Google Drive/blog")
```
```{r shapefiles, message = FALSE}
dt <- read.csv("new_county_election_2016.csv", header=T)
us <- readShapePoly("./USA_adm/USA_adm2.shp")
us0 <- readShapePoly("./USA_adm/USA_adm0.shp")
us.m <- us[-c(which(us$NAME_1=="Alaska")),] #get rid of Alaska
us.d <- us.m[-c(67:71),]
```
5. Prepare color palette
```{r palette, message = FALSE}
nclassint <- 5 #number of colors to be used in the palette
cat.T <- classIntervals(dt$Trump[-c(67:71)], nclassint,style = "jenks") #style refers to how the breaks are created
colpal.T <- brewer.pal(nclassint,"Reds")
color.T <- findColours(cat.T,colpal.T) #sequential
bins.T <- cat.T$brks
lb.T <- length(bins.T)
```
6. Plot the map with map basic...
```{r plot1, message = FALSE}
# pdf("Where are the trump voters.pdf")
# plot(us.d, col=color.T, border=F)
# plot(us0,add=T, lwd=0.1)
# legend("bottomleft",fill=colpal.T,legend=paste(round(bins[-length(bins.T)],1),":",round(bins.T[-1],1)),cex=1, bg="white")
# dev.off()
```
... or ggplot2

```{r plot2, message = FALSE}
library(ggplot2)
library(scales)
theme_set(theme_bw())

ggplot(county_data, aes(Population / Area, Trump)) +
  geom_point() +
  geom_point(data=county_data[which(county_data$State=="Texas"),], aes(x=Population/Area, y=Trump), colour="red")+
  scale_x_log10() +
  scale_y_continuous(labels = percent_format()) +
  xlab("Population density (ppl / square mile)") +
  ylab("% of votes going to Trump") +
  geom_text(aes(label = County), vjust = 1, hjust = 1, check_overlap = TRUE) +
  geom_smooth(method = "lm") +
  ggtitle("Population density vs Trump voters by county (Texas Counties in red)")

# ggplot(county_data, aes(Population / Area, Trump)) +
#   geom_point() +
#   geom_point(data=county_data[which(county_data$State=="California"),], aes(x=Population/Area, y=Trump), colour="red")+
#   scale_x_log10() +
#   scale_y_continuous(labels = percent_format()) +
#   xlab("Population density (ppl / square mile)") +
#   ylab("% of votes going to Trump") +
#   geom_text(data=county_data[which(county_data$State=="California"),], aes(label = ifelse(Trump>.5, as.character(dt$County), "" )), color= "red",size=5,vjust = 1, hjust = 1, check_overlap = TRUE) +
#   geom_smooth(method = "lm") +
#   ggtitle("Population density vs Trump voters by county (California in red)")
```
